Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-22 paper/main.tex
composition

\subsection{Categorical Artist}
We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$. This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once. This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

Since the artist is a morphism of sheafs
\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow Im_A
\end{equation}
it is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}. This means that, by definition, all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. This is the basis of the equivariance illustrated in \autoref{eq:artist:equivariance}

\begin{equation}
  \label{eq:artist:equivariance}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist}                                                 
  \end{tikzcd}
\end{equation}
which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$. This commutativity is described as the matching of data and visual structure by Kindlmann and Scheidegger when introducing their analougous correspondence principal. In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

\subsection{Multiple Graphics: Superposition and Composition}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig:artist:union:glue}}
%\hfil
\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig:artist:union:continuity}}
\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal. The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image. \note{add a border/make the same size/etc so 1 doesn't run over into two}}
\end{figure}
%% artist composition is fiber products/sums of artists on sheafs -> closed cartesian category maybe?

Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels. \note{formalize superposition, than break down what that means for composition}
\subsubsection{Superposition: Multiple Datasets}
\note{far as we care, this can technically be A = A1 + A2....}

\subsubsection{Composite Artist: Multivariate Datasets}
\note{This is A = A_1 \circ A_2 b\c it's sharing nus or xis or all of the above}

\paragraph{Overlapping Base Spaces}

\paragraph{Shared Fibers}

>>>>

\subsection{Categorical Artist}
\label{sec:artist:category}
We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$. This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once. This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}. \note{does this actually go down in the composition section?}

@@

Since the artist is a morphism of sheafs
\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow Im_{\vartist}
\end{equation}
it is by definition a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}. This means that all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. This is the basis of the equivariance condition illustrated in \autoref{eq:artist:equivariance}

\begin{equation}
  \label{eq:artist:equivariance}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist}                                                 
  \end{tikzcd}
\end{equation}
which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$. This commutativity property states that visualization library components must be implemented such that changes in data induce equivalent changes in the output, which is how visualization libraries can produce visualizations that satisfy Kindlmann and Scheidegger's correspondence principal that data and visual structure must match.

In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

\subsection{Composition of Artists}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig:artist:union:glue}}
%\hfil
\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig:artist:union:continuity}}
\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space. The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image. \note{add a border/make the same size/etc so 1 doesn't run over into two}}
\end{figure}

\note{is about building artist that preserves constraint A1 out of A2 + A3}
% generalizing sheaf to collection of sheafs w/ inclusion maps so can tell you how to assemble A1 + A2 such that that push out is preserved 
% when we build an artist out of nus and qs, this is how to preserve this 
% if the artist is a black box, what are the pre and post conditionsg
Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels. Many times, visualizations are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}, and there is an expectation of consistency in how this data is visualized\cite{hullmanKeeping2018}. In \autoref{fig:artist:union:glue}, the disjointedness of the graphic lends itself to an interpretation that the data is disjoint, whereas the consistency of the visual encodings in \autoref{fig:artist:union:continuity} indicates that visual elements share data attributes. In this section, we describe how to express the structure of the data input in terms of  whether the sheaf objects $\sheaf_{\mathcal{\dtotal}}$ are disjoint or share attributes. We then propose that a topologically equivariant artist will preserve this structure in its output.

\subsection{Union of Base Spaces}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{union_k.png}
  \caption{}
  \label{fig:artist:compose:union_base}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{disjoint_k.png}
  \caption{}
  \label{fig:artist:compose:disjoint_base}
\end{figure}

\paragraph{Shared fibers}

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{intersection_f.png}
  \caption{}
  \label{fig:artist:compose:union_fiber}
\end{figure}
